Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-01-22 PAPER.tex
-Some edits to covariance language in Section 2 -Updated Figure B1 y-axis

This paper adds to the growing foundations of lessons which have been documented in  \cite{Patil2016,Jacobs2016} and \cite{Paciga2013} by the LOFAR, MWA and GMRT projects respectively.

@@

There are many choices a $21$ cm data analyst must consider, though the space is significantly constrained by 
the type of instrument being used.  The PAPER instrument in its grid configuration focuses most sensitivity on 
only a few baselines but imaging performance is poor. In this space, signal processing tricks which filter or 
down-weight foreground modes without reference to a sky model have the advantage of being agnostic about 
the sky. However, even within this problem sub-space, questions remain.

 How can time-ordered measurements be combined? How can the 
variance of the data be estimated? In what way(s) can the data be weighted to suppress contaminated modes while not 
destroying an EoR signal? How can a statistically significant detection of a signal be properly identified? Many common techniques, such as 
averaging data, weighting, bootstrapping, and jackknife testing, address these issues but harbor additional trade-offs. For 
example, an aggressive filtering method may succeed in eliminating interfering systematics but comes at the cost of losing 
some EoR signal. A chosen weighting scheme may maximize sensitivity but fail to suppress foregrounds.

@@

The following techniques each have advantages when it comes to maximizing sensitivity and understanding systematics in 
data. However, some have limitations, and we will discuss circumstances in which there are trade-offs. We choose to focus on 
these four techniques because they represent major steps in PAPER's power spectrum pipeline, with several of them also 
being standard steps in general $21$ cm analyses.
\begin{itemize}
\item \textbf{Fringe-rate filtering:} Fringe-rate filtering is an averaging scheme for time-ordered data 
(\citealt{parsons_et_al2016}). Broadly, a fringe-rate filter averages visibilities in time to produce a smaller number of more sensitive 
independent samples. However, by adjusting the shape of the filter to match the rate of sidereal motion it can also affect the presence 
of foregrounds and systematics. We explain the trade-offs of filtering in more detail in Section \ref{sec:toymodel_frf}.
\item \textbf{Weighting:} A dataset can be weighted to emphasize certain features and minimize others. One particular flavor of 
weighting employed by PAPER is inverse covariance weighting in frequency, which is a generalized version of inverse variance 
weighting that also takes into account frequency correlations (\citealt{liu_tegmark2011}; \citealt{dillon_et_al2013a}; \citealt{liu_et_al2014a}; \citealt{liu_et_al2014b}; \citealt{dillon_et_al2014}; \citealt{dillon_et_al2015}). This weighting has the effect of down-weighting correlated 
information (i.e. foregrounds) and up-weighting noise-like information (i.e. EoR). However, a challenge of inverse covariance 
weighting is in accurately defining a covariance matrix that best describes the data. We investigate the impact of different types of weighting on signal loss in Section \ref{sec:SiglossOverview}.
\item \textbf{Bootstrapping:} In addition to using theoretical models for covariance matrices and theoretical error estimation 
methods, bootstrapping is one way to estimate errors. Bootstrapping is a useful method for estimating errors of a dataset from 
itself (\citealt{andrae2010}). By randomly drawing many subsamples of the data, we obtain a sense of its inherent variance, though there are subtleties to 
consider such as the independence of values in a dataset. We explore this potential pitfall of bootstrapping in Section \ref{sec:ErrorOverview}.
\item \textbf{Jackknife testing:} A resampling technique useful for estimating bias, jackknives can be taken along different 
dimensions of a dataset to cross-validate results. In particular, null tests can be used to verify whether results are free of 
systematics, as done with CO power spectra (\citealt{keating_et_al2016}) and CMB measurements (see e.g. \citealt{ade_et_al2008}; \citealt{chiang_et_al2010}; \citealt{bischoff_et_al2011}; \citealt{das_et_al2011}; \citealt{araujo_et_al2012}; \citealt{crites_et_al2015}; \citealt{ade_et_al2016}; \citealt{ade_et_al2017}; \citealt{sherwin_et_al2017}). An EoR detection must pass jackknife and null tests, which we highlight in Section \ref{sec:JackknifeOverview}.
\end{itemize}

@@

\noindent where $\widehat{\textbf{p}}$ is the estimate of the true power spectrum. The data analyst has a choice for $
\textbf{M}$ (subject to the constraint that window functions are sensible) --- for simplicity in this section we choose $\textbf{M}$ to be diagonal, although we explore other cases for the analysis 
of PAPER-64 data as explained in Section \ref{sec:Bias}.

In the next three sections, we use toy models to investigate the effects of weighting matrices on signal loss by experimenting with different matrices $\textbf{R}$ and examining their impact on the resulting power spectra estimate $\widehat{\textbf{p}}$. Our goal in 
experimenting with weighting is to suppress foregrounds and investigate EoR losses associated with it. We note that we 
purposely take a thorough and pedagogical approach to describing the toy model examples given in the next few sections. The specifics of how signal loss appears in PAPER's analysis is later described in Section 
\ref{sec:Sigloss}.

@@

To build our intuition into how a particular choice of weighting, namely inverse covariance weighting, can lead to signal loss, we 
use a toy model. We construct a simple dataset that contains 2-dimensional data (representing visibility data with $100$ time 
integrations and $20$ frequency channels). This model represents realistic dimensions of about an hour of PAPER data which 
might be used for a power spectrum analysis. For PAPER-64 (both the previous analysis and our new analysis) we use $\sim$$8$ hours of data for our analysis (with channel 
widths of $0.5$ MHz and integration times of $43$ seconds), but here we scale it down for this toy model with no loss of 
generality.

We create mock visibilities, $\textbf{x}$, and assume a non-tracking, drift scan observation. Hence, flat spectrum sources (away 
from zenith) oscillate in time and frequency in our measurements. We therefore form a mock bright foreground signal, $
\textbf{x}_{FG}$, as a complex sinusoid that varies smoothly in time and frequency, a simplistic but realistic representation of a 
single bright source. We also create a mock EoR signal, $\textbf{x}_{EoR}$, as a complex, Gaussian random signal. A more 
realistic EoR signal would have a sloped power spectrum in $\textbf{p}$(k) (instead of flat, as in the case of white noise), 
which could be simulated by introducing frequency correlations into the mock EoR signal. However, here we treat all 
$k$'s separately, so a simplistic white noise approximation can be used. Our combined data matrix is then $\textbf{x} = 
\textbf{x}_{FG} + \textbf{x}_{EoR}$, to which we apply different weighting schemes. The three data components are shown in Figure 
\ref{fig:toy_sigloss1}.

@@

One choice for the weighting matrix $\textbf{R}$ is an inverse covariance matrix. This type of weighting is attractive for power spectrum analyses because it yields the smallest possible error bars on a measurement. Said differently, it gives the minimum variance estimate of the power spectrum of a signal (\citealt{tegmark_et_al1997a}; \citealt{bond_et_al1998}). In \citet{liu_tegmark2011}, it was shown that inverse covariance weighting also serves as a way to down-weight foregrounds since foregrounds have large covariances in frequency, motivating our use of the weighting for the PAPER analysis.

The covariance matrix $\textbf{C}$, which describes covariances between frequency channels, can be estimated in a variety of ways. Given perfect foreground, 
instrumental, and EoR models, we could form $\textbf{C}$ in a way that accurately describes our measured data. However, if 
our foreground model is flawed, for example, our estimate of $\textbf{C}$ would not be successful at down-weighting them in 
the data.

Therefore, one attractive way to estimate $\textbf{C}$ is to empirically derive it from the data $\textbf{x}$ itself. For instance, one could time-average such that:

@@

\noindent assuming $\langle\textbf{x}\rangle_{t} = 0$ (a reasonable assumption since fringes average to $0$ over a sufficient 
amount of time), where $\langle \rangle_{t}$ denotes a finite average over time. The weighting matrix for inverse covariance weighting is then $
\textbf{R} = \widehat{\textbf{C}}^{-1}$.

First, we compute the power spectrum of $\textbf{x}$ using OQE formalism and a weighting matrix of $\widehat{\textbf{C}}^{-1}$. 
The result is shown in green in the left plot of Figure \ref{fig:toy_sigloss3}. Also plotted in the figure are the unweighted ($\textbf{R} = \textbf{I}$) power spectrum of $
\textbf{x}_{FG}$ (blue) and $\textbf{x}_{EoR}$ (red).

@@

As shown, our inverse covariance-weighted result successfully suppresses foregrounds. It is also evident that our result fails to 
recover the EoR signal --- it exhibits the correct shape, but the amplitude level is slightly low. This is evidence of signal loss. In 
order to understand the behavior of this result, we can closely study our covariance eigenspectrum, shown in Figure \ref{fig:toy_sigloss2}.
%, which is shown in Figure \ref{fig:toy_sigloss12}.

If $\textbf{C}$ is computed from the data itself, it carries the risk of over-fitting information in the data and introducing a 
multiplicative bias (per $k$) to estimates of the signal. For a toy model mathematical derivation of signal loss arising from a data-estimated covariance matrix, see Appendix \ref{sec:sigloss_appendix}. Here we will describe the origin of this signal loss 
intuitively.

If the covariance is estimated from our data, its eigenspectrum differs from the eigenspectrum of the true covariance $\textbf{C}$, and this difference has consequences on our result. An eigenspectrum ranks the eigenvalues of a matrix from 
highest to lowest and can be thought of as a spectrum of weights that are given to each spectral mode in the data. In other 
words, the eigenvalues encode the strength of different shapes in the dataset. The eigenspectrum of the identity matrix $
\textbf{I}$ is flat (all $1$'s) because it gives equal weighting to all modes. This is usually not the case for a covariance matrix, 
for which a sloped eigenspectrum means that modes are given different weights in OQE power spectrum estimation. The modes with the highest eigenvalues are 
down-weighted the most.

@@

If the true covariance matrix $\textbf{C}$ of our data was known, then every single eigenvalue and eigenvector of $\textbf{C}$ would be 
representative of real fluctuations in the data. However, when using an estimated $\widehat{\textbf{C}}$ that is derived from one 
particular data realization, its eigenvalues and eigenvectors may differ from the truth. Said differently, shapes that may not exist (or exist rarely) in a true covariance may appear stronger in the estimated covariance. Hence, they will be down-weighted 
more than they should be.

In general, the strongest modes of $\widehat{\textbf{C}}$ (highest eigenvalues) are expected to be dominated by bright foregrounds 
--- the most prominent shapes in the data. For these `strong' modes, where foregrounds outshine the EoR signal, $
\widehat{\textbf{C}} \sim \textbf{C}$ and down-weighting these modes is beneficial. This is demonstrated in the toy model by the 
successful suppression of the foreground mode, where our estimated covariance matrix identifies the sinusoid shape and 
assigns it the highest eigenvalue (the first peak in Figure \ref{fig:toy_sigloss2}). In Figure \ref{fig:toy_sigloss12} we show the 
covariances of our toy model datasets along with inverse covariance weighted data. The foreground sinusoid is clearly visible in 
$\widehat{\textbf{C}}_{FG}$.% but effectively absent from $\widehat{\textbf{C}}^{-1}_{FG}\textbf{x}_{FG}$.

@@

The danger of an empirically estimated covariance matrix comes from not being able to describe the `weak' eigenmodes of $
\textbf{C}$ accurately, for which the EoR signal is brighter than foregrounds. In such a case, the empirical covariance believes these modes should be down-weighted more than the true covariance would. For these modes, the error that results from not computing $\textbf{C}$ accurately (due to a finite data sample), has a larger impact on the `weak' modes than the `strong' modes. In other words, mis-estimating $\textbf{C}$ affects EoR-dominated eigenmodes more than FG-dominated modes, and it is primarily the coupling between these mis-estimated modes to the data that leads to signal loss.

@@

In the limit of no instrumental noise, the data that we measure, $\x$, is comprised of two signals, such that
\begin{equation}
\x \equiv \f + \s,
\end{equation}
where $\mathbf{f}$ represents the foregrounds and $\mathbf{s}$ represents the cosmological signal. In general, suppose that our power spectrum algorithm passes the data through some function $g$, yielding a lossy estimate of the power spectrum $\phat_{x}$. This can be parametrized as
\begin{equation}
\label{eq:LinearPspecSum}
\langle \phat_{x} \rangle  = \langle g(\x) \rangle = \ell_{\rm fg} \p_{\rm fg} + \ell_{\rm eor} \p_{\rm eor},
\end{equation}
where $\ell_{\rm eor}$ and $\ell_{\rm fg}$ are multiplicative factors accounting for the signal loss in the true EoR 
power spectrum $\p_{\rm eor}$ and true foreground power spectrum $\p_{\rm fg}$, respectively. It is not 
\emph{a priori} obvious why this parameterization is suitable; we thus provide a toy model in Appendix 
\ref{sec:sigloss_appendix} to motivate this, although it should be noted that the derivation is an approximation 
which assumes that the covariance used in the OQE analysis is close to the true covariance with effects due to 
small sample size limited to first order perturbations and thus neglects cross correlation between $x$ and $e$.  
In actual fact the sample size is similar to the number of independent modes, a case were one expects cross 
terms to be remain significant.

@@

Using this motivation, the numerator of Equation \eqref{eq:Deriv2} is precisely our expression for $P_{\rm out}$ (Equation \eqref{eq:sigloss}), and the denominator is $P_{\rm in}$ (Equation \eqref{eq:Pin}), where the function $g$ is our OQE power spectrum pipeline. \dcj{Here is the post-hoc justification for this signal loss method.  Except goes back to assumption of equation for g which just \emph{assumes} that there are no cross terms in the estimation of a power spectrum. So we're just kicking the can down the road.} \acl{Ok, in my defense, Equation \eqref{eq:LinearPspecSum} is a little better than an \emph{assumption}. Appendix \eqref{sec:sigloss_appendix} does in fact illustrate this with an example. The crucial question is whether we're allowed to motivate our method based on a relationship that is only true in expectation. Note that this expectation includes allowances for signal loss due to the cross correlation between signal and foregrounds, with an example of this being Appendix \ref{sec:sigloss_appendix}.}
\jp{I'm fine with leaving the text as is and calling this note closed with the addition of the appendix, but will wait for sign off from DCJ.}
\dcj{ Assuming that expansion to "leading order in $\eta$" means $\eta<<1$ then the approximation is that the 
covariance is nearly perfectly calculated from the data with only small effects due to finite samples. Thus 
correlations don't make it into the expression and foregrounds remain uncorrelated with eor or injected signals.  
I added words explaining this but it still looks strange for us to be saying that loss is linear in one section and 
then explaining in the next section exactly how cross terms were what messed us up in the first go round.}

Effectively, this means that the relationship between the input and output power spectra, $P_{\rm in}$ and $P_{\rm out}$, can be thought of as a transfer function 
which, for a sampling of $P_{\rm in}$ and $P_{\rm out}$ provides a mapping from an input power spectrum distribution into an output 
distribution. By viewing data through this signal loss lens, we may then ask the question ``what input power spectrum 
distribution could this (signal-loss affected) data come from?"

@@

The treatment we outlined in the previous section implicitly assumed that signal loss corrections could be treated in expectation. In practice, however, the signal loss itself is a random variable --- some realizations of the cosmological signal may be more correlated with foreground modes than others, leading to more signal loss. This leads to two issues. The first is that Equation \eqref{eq:LinearPspecSum} may not hold. In particular, a third term that is a quadratic combination of $\f$ and $\s$ (e.g., $\f^\dagger \C^{-1} \s$) may appear. However, we can circumvent this issue by decomposing $\f$ into a sum of two vectors, one that is proportional to $\s$ and one that is orthogonal to $\s$. The former can be absorbed into the EoR term (part of the signal loss estimate), while the latter can be absorbed into the foreground term. In other words, suppose that we have
\begin{equation}
\phat_x \approx \ell_{\rm fg} |\mathbf{f}|^2 + \ell_{\rm eor} |\mathbf{e}|^2 + \ell_{\rm fe} \mathbf{f} \cdot \mathbf{e}.
\end{equation}
in place of Equation \eqref{eq:LinearPspecSum}. Now, split $\mathbf{f}$ into $\mathbf{ f}_\parallel +\mathbf{ f}_\perp$, which are the components of the foregrounds that couple to the EoR versus those that do not. Further parameterize $\mathbf{ f}_\parallel \equiv \rho \mathbf{e}$, where $\rho$ is some correlation coefficient that quantifies the degree of correlation between the foregrounds and the EoR. This then gives
\begin{equation}
\label{eq:fperpfpara}
\phat_x \approx \ell_{\rm fg} |\mathbf{f}|^2 +( \ell_{\rm eor} + \rho  \ell_{\rm fe} ) |\mathbf{e}|^2 + \ell_{\rm fe} \mathbf{f}\cdot \mathbf{f}_\perp.
\end{equation}
Since $|\mathbf{e}|^2$ would be a lossless estimate of the true EoR power spectrum, this expression recovers the linear relation between the true EoR power spectrum and the lossy estimate.
\dcj{Wait what? I think you are saying essentially that some x can remain in Pout which is what we are seeing. Or is this some extra bit of method which has been added to the analysis...} \acl{I'm not sure that's what I'm saying. (And by ``I'm not sure" I don't mean it in the colloquial sense of ``you're not right and I'm just trying to be polite about it", but rather the literal sense of the phrase ``I'm not sure"). What I was trying to say was that if I wrote Equation \eqref{eq:LinearPspecSum} without the ensemble average, I could say
\begin{equation}
\phat_x \approx \ell_{\rm fg} f^2 + \ell_{\rm eor} e^2 + \ell_{\rm fe} fe.
\end{equation}
(Note that even this is an approximation. In principle there is an infinite series containing all possible powers of 
$e$ and $f$). Now, we split $f$ into $ f_\parallel + f_\perp$, which are the components of the foregrounds that 
couple to the EoR versus those that do not. Further parameterize $f_\parallel \equiv \rho e$, where $\rho$ is 
some correlation coefficient that quantifies the degree of correlation between the foregrounds and the EoR. 
Under this parameterization, one can write the $fe$ term in terms of $e^2$ (which gets absorbed into the other 
EoR term, plus residual foregrounds that we don't care about).}\acl{Just promoted this last part to something in 
the main text.}\dcj{I will admit to being a little lost here.  Doesn't changing the basis just convert the unknown f 
dot s term into other unknown terms. Isn't the whole point of Eq 24 supposed to be that we can subtract off the 
$g(x)$ to estimate $\hat{\ell}_{\rm eor}$?  How do we subtract off $f\dot f_\perp$? Don't we want to know $
\ell_{\rm eor}$ not $(\ell_{\rm eor} + \rho\ell_{\rm fe})$? }

>>>>

@@

There are many choices a $21$ cm data analyst must consider. How can time-ordered measurements be combined? How can the 
variance of the data be estimated? In what way(s) can the data be weighted to suppress contaminated modes while not 
destroying an EoR signal? How can a statistically significant detection of a signal be properly identified? Many common techniques, such as 
averaging data, weighting, bootstrapping, and jackknife testing, address these issues but harbor additional trade-offs. For 
example, an aggressive filtering method may succeed in eliminating interfering systematics but comes at the cost of losing 
some EoR signal. A chosen weighting scheme may maximize sensitivity but fail to suppress foregrounds.

@@

The following techniques each have advantages when it comes to maximizing sensitivity and understanding systematics in 
data. However, some have limitations, and we will discuss circumstances in which there are trade-offs. We choose to focus on 
these four techniques because they represent major steps in PAPER's power spectrum pipeline, with several of them also 
being standard steps in general $21$ cm analyses.
\begin{itemize}
\item \textbf{Fringe-rate filtering:} Fringe-rate filtering is an averaging scheme for time-ordered data 
(\citealt{parsons_et_al2016}). Broadly, a fringe-rate filter increases the sensitivity of a dataset and reduces the number of 
independent samples by an amount dependent on the width of the averaging window. However, it can also affect the presence 
of foregrounds and systematics. We explain the trade-offs of filtering in more detail in Section \ref{sec:toymodel_frf}.
\item \textbf{Weighting:} A dataset can be weighted to emphasize certain features and minimize others. One particular flavor of 
weighting employed by PAPER is inverse covariance weighting in frequency, which is a generalized version of inverse variance 
weighting that also takes into account frequency correlations (\citealt{liu_tegmark2011}; \citealt{dillon_et_al2013a}; \citealt{liu_et_al2014a}; \citealt{liu_et_al2014b}; \citealt{dillon_et_al2014}; \citealt{dillon_et_al2015}). This weighting has the effect of down-weighting correlated 
information (i.e. foregrounds) and up-weighting noise-like information (i.e. EoR). However, a challenge of inverse covariance 
weighting is in estimating a covariance matrix that is closest to the true covariance of the data; the discrepancy between the two has large impacts on signal loss. We investigate the impact of different types of weighting on signal loss in Section \ref{sec:SiglossOverview}.
\item \textbf{Bootstrapping:} In addition to using theoretical models for covariance matrices and theoretical error estimation 
methods, bootstrapping is one way to estimate errors. Bootstrapping is a useful method for estimating errors of a dataset from 
itself (\citealt{andrae2010}). By randomly drawing many subsamples of the data, we obtain a sense of its inherent variance, though there are subtleties to 
consider such as the independence of values in a dataset. We explore this potential pitfall of bootstrapping in Section \ref{sec:ErrorOverview}.
\item \textbf{Jackknife testing:} A resampling technique useful for estimating bias, jackknives can be taken along different 
dimensions of a dataset to cross-validate results. In particular, null tests can be used to verify whether results are free of 
systematics, as done with CO power spectra (\citealt{keating_et_al2016}) and CMB measurements (see e.g. \citealt{ade_et_al2008}; \citealt{chiang_et_al2010}; \citealt{bischoff_et_al2011}; \citealt{das_et_al2011}; \citealt{araujo_et_al2012}; \citealt{crites_et_al2015}; \citealt{ade_et_al2016}; \citealt{ade_et_al2017}; \citealt{sherwin_et_al2017}). An EoR detection must pass jackknife and null tests, which we highlight in Section \ref{sec:JackknifeOverview}.
\end{itemize}

@@

\noindent where $\widehat{\textbf{p}}$ is the estimate of the true power spectrum. The data analyst has a choice for $
\textbf{M}$ (subject to the constraint that window functions are sensible) --- for simplicity in this section we choose $\textbf{M}$ to be diagonal ($\textbf{M} \propto \textbf{I}$), although we explore other cases for the analysis 
of PAPER-64 data as explained in Section \ref{sec:Bias}.

In the next three sections, we use toy models to investigate the effects of weighting matrices on signal loss by experimenting with different matrices $\textbf{R}$ and examining their impact on the resulting power spectrum estimate $\widehat{\textbf{p}}$. Our goal in 
experimenting with weighting is to suppress foregrounds and investigate EoR losses associated with it. We note that we 
purposely take a thorough and pedagogical approach to describing the toy model examples given in the next few sections. The specifics of how signal loss appears in PAPER's analysis is later described in Section 
\ref{sec:Sigloss}.

@@

One choice for the weighting matrix $\textbf{R}$ used in power spectrum analysis is an inverse covariance matrix. This type of weighting is attractive for power spectrum analyses because it yields the smallest possible error bars on a measurement. Said differently, it gives the minimum variance estimate of the power spectrum of a signal (\citealt{tegmark_et_al1997a}; \citealt{bond_et_al1998}). In \citet{liu_tegmark2011}, it was shown that inverse covariance weighting also serves as a way to down-weight foregrounds since foregrounds have large covariances in frequency, motivating our use of the weighting for previous PAPER analyses.

One important feature of weighting data by its true inverse covariance is that, while it can suppress some foregrounds, it cannot suppress the EoR signal (nor foreground modes that masquerade as signal-like modes). By definition, inverse covariance weighting does not lead to signal loss. Therefore, in an ideal world with perfect foreground, 
instrumental, and EoR models, we could form $\textbf{C}$ in a way that accurately describes our measured data and use inverse covariance weighting to produce minimum variance power spectrum estimates without destroying the signal.

In practice, we do not have perfect models for $\textbf{C}$, and it is the use of an estimated covariance $\widehat{\textbf{C}}$ instead of the true covariance $\textbf{C}$ that can lead to loss. We will now build intuition into how estimated inverse covariance weighting can give rise to signal loss through the use of a toy model.

We construct a simple dataset that contains 2-dimensional data (representing visibility data with $100$ time 
integrations and $20$ frequency channels). This model represents realistic dimensions of about an hour of PAPER data which 
might be used for a power spectrum analysis. For PAPER-64 (both the previous analysis and our new analysis) we use $\sim$$8$ hours of data for our analysis (with channel 
widths of $0.5$ MHz and integration times of $43$ seconds), but here we scale it down for this toy model with no loss of 
generality.

We create mock visibilities, $\textbf{x}$, and assume a non-tracking, drift scan observation. Hence, flat spectrum sources (away 
from zenith) oscillate in time and frequency in our measurements. We therefore form a mock bright foreground signal, $
\textbf{x}_{FG}$, as a complex sinusoid that varies smoothly in time and frequency, a simplistic but realistic representation of a 
single bright source. We also create a mock EoR signal, $\textbf{x}_{EoR}$, as a complex, Gaussian random signal. A more 
realistic EoR signal would have a sloped power spectrum in $\textbf{p}$(k) (instead of flat, as in the case of white noise), 
which could be simulated by introducing frequency correlations into the mock EoR signal. However, here we treat all 
$k$'s separately, so a simplistic white noise approximation can be used. Our combined data matrix is then $\textbf{x} = 
\textbf{x}_{FG} + \textbf{x}_{EoR}$, to which we apply different weighting schemes throughout Section \ref{sec:SiglossOverview}. The three data components are shown in Figure 
\ref{fig:toy_sigloss1}.

@@

Without a perfect model for the covariance matrix $\textbf{C}$ of our data, one attractive way to estimate it is to empirically derive it from the data $\textbf{x}$ itself (this is what was done in the previous PAPER analyses). For instance, one could time-average such that:

@@

\noindent assuming $\langle\textbf{x}\rangle_{t} = 0$ (a reasonable assumption since fringes average to $0$ over a sufficient 
amount of time), where $\langle \rangle_{t}$ denotes a finite average over time. The weighting matrix for estimated inverse covariance weighting is then $
\textbf{R} = \widehat{\textbf{C}}^{-1}$.

First, we compute the power spectrum of our toy model dataset $\textbf{x}$ using OQE formalism and $\textbf{R} = \widehat{\textbf{C}}^{-1}$. 
The result is shown in green in the left plot of Figure \ref{fig:toy_sigloss3}. Also plotted in the figure are the unweighted ($\textbf{R} = \textbf{I}$) power spectrum of $
\textbf{x}_{FG}$ (blue) and $\textbf{x}_{EoR}$ (red).

@@

As shown, our $\widehat{\textbf{C}}^{-1}$ weighted result successfully suppresses foregrounds. It is also evident that our result fails to 
recover the EoR signal --- it exhibits the correct shape, but the amplitude level is slightly low. This is evidence of signal loss. In short, if the covariance is computed from the data itself, it carries the risk of over-fitting information in the data and introducing a 
multiplicative bias (per $k$) to estimates of the signal. For a toy model mathematical derivation of signal loss arising from a data-estimated covariance matrix, see Appendix \ref{sec:sigloss_appendix}. Here we will describe the origin of this signal loss 
intuitively.

To begin to understand the lossy behavior of this result, we can closely study our covariance eigenspectrum, shown in Figure \ref{fig:toy_sigloss2}. If the covariance is estimated from our data, its eigenspectrum differs from the eigenspectrum of the true covariance $\textbf{C}$, and this difference has consequences on our result. An eigenspectrum ranks the eigenvalues of a matrix from 
highest to lowest and can be thought of as a spectrum of weights that are given to each spectral mode in the data. In other 
words, the eigenvalues encode the strength of different shapes in the dataset. The eigenspectrum of the identity matrix $
\textbf{I}$ is flat (all $1$'s) because it gives equal weighting to all modes. This is usually not the case for a covariance matrix, 
for which a sloped eigenspectrum means that modes are given different weights in OQE power spectrum estimation. The modes with the highest eigenvalues are 
down-weighted the most.

@@

Because $\widehat{\textbf{C}}$ is estimated from the data, its eigenvectors and eigenvalues are strongly coupled to the particular data realization that was used to compute it. The strongest mode of $\widehat{\textbf{C}}$ (highest eigenvalue) describes the sinusoid foreground mode in the toy model (the first peak in Figure \ref{fig:toy_sigloss2}). In Figure \ref{fig:toy_sigloss12} we show the covariances of our toy model datasets along with inverse covariance weighted data. The foreground sinusoid is clearly visible in $\widehat{\textbf{C}}_{FG}$.

In general, the strongest eigenmodes of $\widehat{\textbf{C}}$ typically describe bright foregrounds --- the most prominent shapes in a dataset. For these `strong' modes, where foregrounds outshine the EoR signal, down-weighting is beneficial. In some sense, we desire signal loss, if by `signal' we mean `foregrounds.' In this case it is okay for the `strong' eigenmodes to be coupled to the data in a way that produces loss. For our toy model, the successful suppression of the foreground mode is demonstrated in Figure \ref{fig:toy_sigloss2} by the missing foreground peak in the weighted power spectrum estimate (green).

%If the true covariance matrix $\textbf{C}$ of our data was known, then every single eigenvalue and eigenvector of $\textbf{C}$ would be 
%representative of real fluctuations in the data. However, when using an estimated $\widehat{\textbf{C}}$ that is derived from one 
%particular data realization, its eigenvalues and eigenvectors may differ from the truth. Said differently, shapes that may not exist (or exist rarely) in a true covariance may appear stronger in the estimated covariance. Hence, they will be down-weighted 
%more than they should be.

@@

The danger of an empirically estimated covariance matrix comes mostly from not being able to describe the `weak' eigenmodes of $
\textbf{C}$ accurately, for which the EoR signal is brighter than foregrounds. \cc{start editing here later!} In such a case, the empirical covariance believes these modes should be down-weighted more than the true covariance would. For these modes, the error that results from not computing $\textbf{C}$ accurately (due to a finite data sample), has a larger impact on the `weak' modes than the `strong' modes. In other words, mis-estimating $\textbf{C}$ affects EoR-dominated eigenmodes more than FG-dominated modes, and it is primarily the coupling between these mis-estimated modes to the data that leads to signal loss.

@@

In the limit of no instrumental noise, the data that we measure, $\x$, is comprised of two signals, such that
\begin{equation}
\x \equiv \f + \s,
\end{equation}
where $\mathbf{f}$ represents the foregrounds and $\mathbf{s}$ represents the cosmological signal. In general, suppose that our power spectrum algorithm passes the data through some function $g$, yielding a lossy estimate of the power spectrum $\phat_{x}$. This can be parametrized as
\begin{equation}
\label{eq:LinearPspecSum}
\langle \phat_{x} \rangle  = \langle g(\x) \rangle = \ell_{\rm fg} \p_{\rm fg} + \ell_{\rm eor} \p_{\rm eor},
\end{equation}
where $\ell_{\rm eor}$ and $\ell_{\rm fg}$ are multiplicative factors accounting for the signal loss in the true EoR power spectrum $\p_{\rm eor}$ and true foreground power spectrum $\p_{\rm fg}$, respectively. It is not \emph{a priori} obvious why this parameterization is suitable; we thus provide a toy model in Appendix \ref{sec:sigloss_appendix} to motivate this, although it should be noted that the derivation is an approximation.

@@

Using this motivation, the numerator of Equation \eqref{eq:Deriv2} is precisely our expression for $P_{\rm out}$ (Equation \eqref{eq:sigloss}), and the denominator is $P_{\rm in}$ (Equation \eqref{eq:Pin}), where the function $g$ is our OQE power spectrum pipeline. \dcj{Here is the post-hoc justification for this signal loss method.  Except goes back to assumption of equation for g which just \emph{assumes} that there are no cross terms in the estimation of a power spectrum. So we're just kicking the can down the road.} \acl{Ok, in my defense, Equation \eqref{eq:LinearPspecSum} is a little better than an \emph{assumption}. Appendix \eqref{sec:sigloss_appendix} does in fact illustrate this with an example. The crucial question is whether we're allowed to motivate our method based on a relationship that is only true in expectation. Note that this expectation includes allowances for signal loss due to the cross correlation between signal and foregrounds, with an example of this being Appendix \ref{sec:sigloss_appendix}.}
\jp{I'm fine with leaving the text as is and calling this note closed with the addition of the appendix, but will wait for sign off from DCJ.}
Effectively, this means that the relationship between the input and output power spectra, $P_{\rm in}$ and $P_{\rm out}$, can be thought of as a transfer function 
which, for a sampling of $P_{\rm in}$ and $P_{\rm out}$ provides a mapping from an input power spectrum distribution into an output 
distribution. By viewing data through this signal loss lens, we may then ask the question ``what input power spectrum 
distribution could this (signal-loss affected) data come from?"

@@

The treatment we outlined in the previous section implicitly assumed that signal loss corrections could be treated in expectation. In practice, however, the signal loss itself is a random variable --- some realizations of the cosmological signal may be more correlated with foreground modes than others, leading to more signal loss. This leads to two issues. The first is that Equation \eqref{eq:LinearPspecSum} may not hold. In particular, a third term that is a quadratic combination of $\f$ and $\s$ (e.g., $\f^\dagger \C^{-1} \s$) may appear. However, we can circumvent this issue by decomposing $\f$ into a sum of two vectors, one that is proportional to $\s$ and one that is orthogonal to $\s$. The former can be absorbed into the EoR term (part of the signal loss estimate), while the latter can be absorbed into the foreground term. In other words, suppose that we have
\begin{equation}
\phat_x \approx \ell_{\rm fg} |\mathbf{f}|^2 + \ell_{\rm eor} |\mathbf{e}|^2 + \ell_{\rm fe} \mathbf{f} \cdot \mathbf{e}.
\end{equation}
in place of Equation \eqref{eq:LinearPspecSum}. Now, split $\mathbf{f}$ into $\mathbf{ f}_\parallel +\mathbf{ f}_\perp$, which are the components of the foregrounds that couple to the EoR versus those that do not. Further parameterize $\mathbf{ f}_\parallel \equiv \rho \mathbf{e}$, where $\rho$ is some correlation coefficient that quantifies the degree of correlation between the foregrounds and the EoR. This then gives
\begin{equation}
\label{eq:fperpfpara}
\phat_x \approx \ell_{\rm fg} |\mathbf{f}|^2 +( \ell_{\rm eor} + \rho  \ell_{\rm fe} ) |\mathbf{e}|^2 + \ell_{\rm fe} \mathbf{f}\cdot \mathbf{f}_\perp.
\end{equation}
Since $|\mathbf{e}|^2$ would be a lossless estimate of the true EoR power spectrum, this expression recovers the linear relation between the true EoR power spectrum and the lossy estimate.
\dcj{Wait what? I think you are saying essentially that some x can remain in Pout which is what we are seeing. Or is this some extra bit of method which has been added to the analysis...} \acl{I'm not sure that's what I'm saying. (And by ``I'm not sure" I don't mean it in the colloquial sense of ``you're not right and I'm just trying to be polite about it", but rather the literal sense of the phrase ``I'm not sure"). What I was trying to say was that if I wrote Equation \eqref{eq:LinearPspecSum} without the ensemble average, I could say
\begin{equation}
\phat_x \approx \ell_{\rm fg} f^2 + \ell_{\rm eor} e^2 + \ell_{\rm fe} fe.
\end{equation}
(Note that even this is an approximation. In principle there is an infinite series containing all possible powers of $e$ and $f$). Now, we split $f$ into $ f_\parallel + f_\perp$, which are the components of the foregrounds that couple to the EoR versus those that do not. Further parameterize $f_\parallel \equiv \rho e$, where $\rho$ is some correlation coefficient that quantifies the degree of correlation between the foregrounds and the EoR. Under this parameterization, one can write the $fe$ term in terms of $e^2$ (which gets absorbed into the other EoR term, plus residual foregrounds that we don't care about).}\acl{Just promoted this last part to something in the main text.}
